---
title: "Time Series Analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: ["twitter", "facebook", "menu"]
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(rpivotTable)
library(openintro)
library(highcharter)
library(ggplot2)
library(plotly)
library(tidyverse)
library(Rcpp)
library(rlang)
library(echarts4r)
library(prophet)
library(gplots)
```

```{r, include=FALSE}
# Setting the environment and Importing data
getwd()
setwd("C:/Users/eseno/Desktop/Projects/R/Energy Project")
data <- read.csv("pjm_hourly_est.csv")
```

```{r, include=FALSE}
#First Impression About Data
dim(data) # very first impression about dimensions of data
str(data) # very first impression about data
data$Datetime <- as.Date(data$Datetime) # Converting dates from chr into date structure 
max(data$Datetime) # Max and min dates to understand the range of data
min(data$Datetime)
data_new <- separate(data,Datetime, c("y","m","d"))## Separating Date into Year, Month and Day
dataRange <- max(data$Datetime) - min(data$Datetime)
colnames(data)
```

First Impression 
=======================================================================
Row {data-height=300, data-weight= 200}
-----------------------------------------------------------------------
### What to Analyse?

```{r}
dataName <- c("PJM Energy Consumption Data")
valueBox(dataName,color = "maroon",icon = "fa-info")
```

### Where to Find the Data Set?

```{r}
dataWhere <- c("Kaggle Open Datasets")
valueBox(dataWhere,color = "warning",icon = "fa-info")
```

Row {data-height=200, data-weight=200}
-----------------------------------------------------------------------
### Available Data (in total)

```{r}
dataRange <- cat("From: Apr. 1998\nTo: Aug. 2018")
valueBox(dataRange,color = "primary",icon = "fa-info")
```

### Number of Observations (in total)

```{r}
numberOfObservations <- nrow(data)
valueBox(numberOfObservations,color = "orange",icon = "fa-calculator")
```

### Number of Columns

```{r}
numberOfColumns <- dim(data)[2]
valueBox(numberOfColumns,color = "green",icon = "fa-calculator")
```

### The Analyzed Data  

```{r}
dataRange <- cat("From: Jan. 2002\nTo: Aug. 2018")
valueBox(dataRange,color = "red",icon = "fa-info")
```

### Number of Observations (included into the analysis)

```{r}
numberOfObservations <- nrow(data) - as.integer(last(sum(is.na(data$PJME))-1))
valueBox(numberOfObservations,color = "gray",icon = "fa-calculator")
```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, include=FALSE}
# Defining a function for plotting graphs
function_graph=function(data,a,b){
  x <- ggplot(data,aes(x=a,y=b))+
    geom_line(color="steel blue")+
    xlab("")+
    theme(panel.background = element_rect(fill = "transparent"),
          plot.background = element_rect(fill = "transparent", color = NA),
          axis.title.x=element_text(colour="Red", size=15),
          axis.text.x=element_text(size=10),
          axis.title.y=element_text(color ="Red", size=15),
          axis.text.y=element_text(size=10),
          plot.title= element_text(size=20,
                                   colour ="dark blue",
                                   hjust = 0.5))}
```

### PJM
``` {r, include = FALSE}
# Finding the breakpoints where NA values start
arrange(data, desc(Datetime))
lastE<- last(sum(is.na(data$PJME)))
lastE
lastW<- last(sum(is.na(data$PJMW)))
lastW
```

``` {r}
# Plotting PJM (total) 
msg <- "Explanation: \nAs we can see in the graph, the data for PJM is only available till 01.01.2002.  \nFrom this point on, the data is available for PJME and PJMW after 01.04.2002. \nSince the data for PJME and PJMW provide more detailled information,\nthis analysis focuses on both in order not to digress from the subject."
function_graph(data=data, a=data$Datetime, b=data$PJM_Load)+
  ylab("PJM Load")+
  ggtitle("PJM Load (Apr. 1998 - Aug. 2018)")+
    labs(subtitle = msg) +
    theme(plot.margin = unit(c(0.5,0.4,0.25,0.25),"cm"),
          rect = element_rect(fill = "transparent",color = NA),
          plot.subtitle = element_text(hjust=0.5,
                                       size = 10,
                                       color = "maroon",
                                       face = "bold"))
```

### PJME

``` {r}
# Plotting PJME Consumption
function_graph(data=data, a=data$Datetime, b=data$PJME)+
  ylab("PJME Load")+
  ggtitle("PJME Load (Apr. 1998 - Aug. 2018)")
  ## an alternative to limit date: scale_x_date(limit = c(as.Date("2015-01-01"),     as.Date("2018-01-01")))
```

### PJMW

``` {r}
# Plotting PJMW Consumption
function_graph(data=data, a=data$Datetime, b=data$PJMW)+
  ylab("PJMW Load")+
  ggtitle("PJMW Load (Apr. 1998 - Aug. 2018)")
``` 


Data
=======================================================================

``` {r, include = FALSE}
# Due to the amount of data being analyzed, I just taken the data in 2018 into consideration in this part.

dataDT <- filter(data, Datetime >= "2018-01-01")%>% 
  filter(Datetime <= "2018-08-03")%>% 
  arrange(Datetime) %>% 
  select(c(1:9,11:12)) #the columns providing only NA values were excluded from the table.
```

**Note**: Due to the amount of the data being analyzed, the data only in 2018 is taken here into consideration in order not to overload the table. 
``` {r}
datatable(data = dataDT,
          caption = "PJM Energy Consumption Data (2018)",
          rownames =  TRUE,
          options = list(pageLegnth = 50),
          filter = "top")
``` 

Missing Value Analysis
=======================================================================

```{r, include=FALSE}
### Missing Value Analysis
na <- apply(data, 2, function(missing_value) any(is.na(missing_value) | is.infinite(missing_value) | is.nan(missing_value)))
na # gives the columns containing NA, Nan and Inf Values
sapply(data, function(x) sum(is.na(x))) # gives the amount of na values in each column
summary(data) # provides a general understanding of data structure 

### A Detailed Analysis of Missing Values
detectingMissingValues_min=function(data){ #detecting where the first non missing value is
  NonNAindex <- which(!is.na(data))
  firstNonNA <- min(NonNAindex)
}
detectingMissingValues_max=function(data){ #detecting where the last non missing value is
  NonNAindex <- which(!is.na(data))
  lastNonNa <- max(NonNAindex)
}

l <- list()
percentageList <- list()
for(i in names(data)){ # detects the numbers of the first and last non missing value in columns 
  Na_i_min <- detectingMissingValues_min(data[i])
  Na_i_max <- detectingMissingValues_max(data[i])
  percentageNA <- round(mean(is.na(data[i])), digits =2)
  index <- tibble(i,Na_i_min,Na_i_max,percentageNA)
  l <- c(l,index)
  percentageList <- c(percentageList,percentageNA)
}
l <- as.matrix(l)
colnames(l) <- c("Values")
print(l) # gives a summary of exact locations of the non missing values in each column 
percentageList <- as.matrix(percentageList)
percentageList # gives a matrix with the amount of missing values and necessary inputs for plotting gauges
```


Row {data-height=650}
-----------------------------------------------------------------------
### **Missing Value Percentages(Power Consumption)**

```{r}
e_charts() %>% 
 e_gauge(18, "PJME", center = c("20%", "50%"), radius = "75%") %>% 
 e_gauge(20, "PJMW", center = c("50%", "50%"), radius = "75%") %>% 
 e_gauge(82, "PJM_Load", center = c("80%", "50%"), radius = "75%")

```

Row {data-height=750}
-----------------------------------------------------------------------

### **Missing Value Percentages(Power Suppliers)**

```{r}
e_charts() %>% 
  e_gauge(32, "AEP", center = c("10%", "30%"), radius = "55%") %>%
  e_gauge(63, "COMED", center = c("30%", "30%"), radius = "55%")%>%
  e_gauge(32, "DAYTON", center = c("50%", "30%"), radius = "55%") %>% 
  e_gauge(68, "DEOK", center = c("70%", "30%"), radius = "55%") %>% 
  e_gauge(35, "DOM", center = c("90%", "30%"), radius = "55%") %>% 
  e_gauge(33, "DUQ", center = c("20%", "80%"), radius = "55%") %>% 
  e_gauge(75, "EKPC", center = c("40%", "80%"), radius = "55%") %>% 
  e_gauge(65, "FE", center = c("60%", "80%"), radius = "55%") %>% 
  e_gauge(67, "NI", center = c("80%", "80%"), radius = "55%")  
```

Detailed Analysis (PJME)
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, include = FALSE}
# Data Cleaning

dataCleaned <- filter(data, Datetime >= "2002-01-02" )%>% # excluding data from analysis before 2002 due to the non-availability of the data
  arrange(Datetime)%>%  
  select(c(-13)) 

year <- separate(dataCleaned, Datetime, c("y"))%>% 
  select(1)

dataCleaned <- cbind(year, dataCleaned)
dataCleaned$y <- as.integer(dataCleaned$y)
str(dataCleaned)

# Subseting Data for each year individually
subset2002 <- subset(dataCleaned, y==2002)
subset2003 <- subset(dataCleaned, y==2003)
subset2004 <- subset(dataCleaned, y==2004)
subset2005 <- subset(dataCleaned, y==2005)
subset2006 <- subset(dataCleaned, y==2006)
subset2007 <- subset(dataCleaned, y==2007)
subset2008 <- subset(dataCleaned, y==2008)
subset2009 <- subset(dataCleaned, y==2009)
subset2010 <- subset(dataCleaned, y==2010)
subset2011 <- subset(dataCleaned, y==2011)
subset2012 <- subset(dataCleaned, y==2012)
subset2013 <- subset(dataCleaned, y==2013)
subset2014 <- subset(dataCleaned, y==2014)
subset2015 <- subset(dataCleaned, y==2015)
subset2016 <- subset(dataCleaned, y==2016)
subset2017 <- subset(dataCleaned, y==2017)
subset2018 <- subset(dataCleaned, y==2018)

# Subsetting Data as Time Intervals
subset2005_TI <-filter(dataCleaned, Datetime >= "2002-01-01")%>% 
             filter(Datetime <= "2005-12-31")
subset2009_TI <-filter(dataCleaned, Datetime >= "2006-01-01")%>% 
             filter(Datetime <= "2009-12-31")
subset2013_TI <-filter(dataCleaned, Datetime >= "2010-01-01")%>% 
             filter(Datetime <= "2013-12-31")
subset2018_TI <-filter(dataCleaned, Datetime >= "2014-01-01")%>% 
             filter(Datetime <= "2018-12-31")
```

```{r, include = FALSE} 
# Defining a function to plot graphs individually

graphAnalysisE=function(data,a,b){  
x <- ggplot(data)+
    geom_line(aes(x=a,y=b),color="steel blue", alpha = 0.7)+
    xlab("")+ ylab("")+
    geom_smooth(aes(x=Datetime, y=PJME), colour = "maroon")+
    stat_smooth(aes(x=Datetime, y=PJME), method = 'lm', se = FALSE, colour =  "magenta1") +
    theme(panel.background = element_rect(fill = "transparent"),
          plot.background = element_rect(fill = "transparent", color = NA),
          axis.title.x=element_text(colour="Red", size=15),
          axis.text.x=element_text(size=10),
          axis.title.y=element_text(color ="Red", size=15),
          axis.text.y=element_text(size=10),
          plot.title= element_text(size=20,
                                   colour ="dark blue",
                                   hjust = 0.5))}

```

### 2002-2005 (PJME)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisE(data=subset2002,a=subset2002$Datetime, b=subset2002$PJME)+
  ggtitle("PJME Consumption (2002)")


graphAnalysisE(data=subset2003,a=subset2003$Datetime, b=subset2003$PJME)+
  ggtitle("PJME Consumption (2003)")

graphAnalysisE(data=subset2004,a=subset2004$Datetime, b=subset2004$PJME)+
  ggtitle("PJME Consumption (2004)")

graphAnalysisE(data=subset2005,a=subset2005$Datetime, b=subset2005$PJME)+
  ggtitle("PJME Consumption (2005)")

```

### 2006-2009 (PJME)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisE(data=subset2006,a=subset2006$Datetime, b=subset2006$PJME)+
  ggtitle("PJME Consumption (2006)")

graphAnalysisE(data=subset2007,a=subset2007$Datetime, b=subset2007$PJME)+
  ggtitle("PJME Consumption (2007)")

graphAnalysisE(data=subset2008,a=subset2008$Datetime, b=subset2008$PJME)+
  ggtitle("PJME Consumption (2008)")

graphAnalysisE(data=subset2009,a=subset2009$Datetime, b=subset2009$PJME)+
  ggtitle("PJME Consumption (2009)")

```

### 2010-2013 (PJME)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisE(data=subset2010,a=subset2010$Datetime, b=subset2010$PJME)+
  ggtitle("PJME Consumption (2010)")

graphAnalysisE(data=subset2011,a=subset2011$Datetime, b=subset2011$PJME)+
  ggtitle("PJME Consumption (2011)")

graphAnalysisE(data=subset2012,a=subset2012$Datetime, b=subset2012$PJME)+
  ggtitle("PJME Consumption (2012)")

graphAnalysisE(data=subset2013,a=subset2013$Datetime, b=subset2013$PJME)+
  ggtitle("PJME Consumption (2013)")

```

### 2014-2018 (PJME)
```{r, fig.show="hold", out.width="30%"}

graphAnalysisE(data=subset2014,a=subset2014$Datetime, b=subset2014$PJME)+
  ggtitle("PJME Consumption (2014)")

graphAnalysisE(data=subset2015,a=subset2015$Datetime, b=subset2015$PJME)+
  ggtitle("PJME Consumption (2015)")

graphAnalysisE(data=subset2016,a=subset2016$Datetime, b=subset2016$PJME)+
  ggtitle("PJME Consumption (2016)")

graphAnalysisE(data=subset2017,a=subset2017$Datetime, b=subset2017$PJME)+
  ggtitle("PJME Consumption (2017)")

graphAnalysisE(data=subset2018,a=subset2018$Datetime, b=subset2018$PJME)+
  ggtitle("PJME Consumption (2018)")
```

### PJME (2002-2005)
```{r}
ggplot(data = subset2005_TI)+
  geom_line(aes(x=Datetime, y=PJME),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJME), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJME (2006-2009)
```{r}

ggplot(data = subset2009_TI)+
  geom_line(aes(x=Datetime, y=PJME),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJME), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJME (2010-2013)
```{r}
ggplot(data = subset2013_TI)+
  geom_line(aes(x=Datetime, y=PJME),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJME), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJME (2014-2018)
```{r}
ggplot(data = subset2018_TI)+
  geom_line(aes(x=Datetime, y=PJME),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJME), colour = "red", fill =NA)+
  facet_grid(y~., scales ="free", space = "free")+
  xlab("")
```

Detailed Analysis (PJMW)
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------
```{r, include=FALSE}
graphAnalysisW=function(data,a,b){  
x <- ggplot(data)+
    geom_line(aes(x=a,y=b),color="steel blue", alpha = 0.7)+
    xlab("")+ ylab("")+
    geom_smooth(aes(x=Datetime, y=PJMW), colour = "maroon")+
    stat_smooth(aes(x=Datetime, y=PJMW), method = 'lm', se = FALSE, colour =  "magenta1") +
    theme(panel.background = element_rect(fill = "transparent"),
          plot.background = element_rect(fill = "transparent", color = NA),
          axis.title.x=element_text(colour="Red", size=15),
          axis.text.x=element_text(size=10),
          axis.title.y=element_text(color ="Red", size=15),
          axis.text.y=element_text(size=10),
          plot.title= element_text(size=20,
                                   colour ="dark blue",
                                   hjust = 0.5))}
```

### 2002-2005 (PJMW)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisW(data=subset2002,a=subset2002$Datetime, b=subset2002$PJMW)+
  ggtitle("PJMW Consumption (2002)")

graphAnalysisW(data=subset2003,a=subset2003$Datetime, b=subset2003$PJMW)+
  ggtitle("PJMW Consumption (2003)")

graphAnalysisW(data=subset2004,a=subset2004$Datetime, b=subset2004$PJMW)+
  ggtitle("PJMW Consumption (2004)")

graphAnalysisW(data=subset2005,a=subset2005$Datetime, b=subset2005$PJMW)+
  ggtitle("PJMW Consumption (2005)")

```

### 2006-2009 (PJMW)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisW(data=subset2006,a=subset2006$Datetime, b=subset2006$PJMW)+
  ggtitle("PJMW Consumption (2006)")

graphAnalysisW(data=subset2007,a=subset2007$Datetime, b=subset2007$PJMW)+
  ggtitle("PJMW Consumption (2007)")

graphAnalysisW(data=subset2008,a=subset2008$Datetime, b=subset2008$PJMW)+
  ggtitle("PJMW Consumption (2008)")

graphAnalysisW(data=subset2009,a=subset2009$Datetime, b=subset2009$PJMW)+
  ggtitle("PJMW Consumption (2009)")

```

### 2010-2013 (PJMW)
```{r, fig.show="hold", out.width="50%"}

graphAnalysisW(data=subset2010,a=subset2010$Datetime, b=subset2010$PJMW)+
  ggtitle("PJMW Consumption (2010)")

graphAnalysisW(data=subset2011,a=subset2011$Datetime, b=subset2011$PJMW)+
  ggtitle("PJMW Consumption (2011)")

graphAnalysisW(data=subset2012,a=subset2012$Datetime, b=subset2012$PJMW)+
  ggtitle("PJMW Consumption (2012)")

graphAnalysisW(data=subset2013,a=subset2013$Datetime, b=subset2013$PJMW)+
  ggtitle("PJMW Consumption (2013)")

```

### 2014-2018 (PJMW)
```{r, fig.show="hold", out.width="30%"}

graphAnalysisW(data=subset2014,a=subset2014$Datetime, b=subset2014$PJMW)+
  ggtitle("PJMW Consumption (2014)")

graphAnalysisW(data=subset2015,a=subset2015$Datetime, b=subset2015$PJMW)+
  ggtitle("PJMW Consumption (2015)")

graphAnalysisW(data=subset2016,a=subset2016$Datetime, b=subset2016$PJMW)+
  ggtitle("PJMW Consumption (2016)")

graphAnalysisW(data=subset2017,a=subset2017$Datetime, b=subset2017$PJMW)+
  ggtitle("PJMW Consumption (2017)")

graphAnalysisW(data=subset2018,a=subset2018$Datetime, b=subset2018$PJMW)+
  ggtitle("PJMW Consumption (2018)")
```

### PJMW (2002-2005)
```{r}
ggplot(data = subset2005_TI)+
  geom_line(aes(x=Datetime, y=PJMW),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJMW), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJMW (2006-2009)
```{r}

ggplot(data = subset2009_TI)+
  geom_line(aes(x=Datetime, y=PJMW),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJMW), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJMW (2010-2013)
```{r}
ggplot(data = subset2013_TI)+
  geom_line(aes(x=Datetime, y=PJMW),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJMW), colour = "red", fill =NA)+
  facet_grid(y~., scales="free")+
  xlab("")
```

### PJMW (2014-2018)
```{r}
ggplot(data = subset2018_TI)+
  geom_line(aes(x=Datetime, y=PJMW),colour ="steel blue",alpha=0.7)+
  geom_smooth(aes(x=Datetime, y=PJMW), colour = "red", fill =NA)+
  facet_grid(y~., scales ="free", space = "free")+
  xlab("")
```

Predictions
=======================================================================



Pivot Table
=======================================================================




